Working document to try to explain why/how ATAs influence coexistence in an intuitive way for referee 3

### 2023-07-03 paper/RefereeResponses1.tex

\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
such errors. But we're happy to hear specific suggestions.} \\

\noindent \textbf{Comments of referee 3:} \\

\noindent C) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked. ATAs can give additional mechanisms that either promote or hinder coexistence in a similar fashion and magnitude to other previously recognised mechanisms such as the storage effect. To illustrate these ideas the authors combine results from mathematical models, simulations, and empirical examples, and show that under a certain combination of parameters ATAs can be a significant mechanism of species coexistence. \\

\noindent ***\emph{Many thanks for the positive feedback!} \\

@@

\noindent C) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations. Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics. I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript. \\

\noindent ***\emph{We have addressed the referee's specific suggestions below. We have also redoubled our
efforts to make the manuscript as accessible as possible.} \\
%We have to redouble our efforts, as it says.

@@

\noindent C) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences. \\

\noindent ***\emph{response} \\
%Jasmin, let's discuss. I cannot, myself, say, even now, precisely *how* or *why* ATAs promote or hinder coexistence. 
%In my view, the paper is about showing that they CAN promote or hinder coexistence, and to what extent, so that 
%future work can continue to explore when they do and when they don't, and precisely how. So I want to write a response
%to that effect. However, perhaps you have an intuition, which can be simply explained, as to how ATAs
%either promote or hinder coexistence. I know that's maybe a lot to ask, but if we have such a thing we should consider
%putting it in. Ecologists always want a two-sentence, math-less explanation for everything. If quantum mechanics were
%part of ecology they'd want a two-sentence intuitive explanation for it, even if it was meaningless aside making
%the reader feel better about themselves. Sometimes such a thing does not exist, but when it does exist it can,
%at least sometimes, be useful. For the record, I think most intuitive, verbal explanations of storage effects
%are mostly useless, but perhaps we can do better than those.

@@

\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, but how ATAs enter in play is not well solved. It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented. \\

\noindent ***\emph{response} \\

\noindent C) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor. Could the authors briefly indicate how this could occur? \\

\noindent ***\emph{<It cannot. ATAs can only influence storage effects, not the other mechanisms, as explained on 
lines xyz. Is that enough?>} \\

@@

\noindent C) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\

\noindent ***\emph{<This is a good question. We need to find a place where the models are presented to motivate
them, AND a place later, in the Discussion (or a new SI section called ``Additional discussion points'') where
we talk about why ATAs matter for one but not the other. I think this is just because the
effects of extremes are accentuated in the beta one, and so the effects of associations in the extremes are
accentuated. But that's not unreasonable, since many species have environments which fluctuate between
being suitable and unsuitable.>} \\

@@

\noindent C) The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model. \\

%This referee is asking us to explain a lot of things in more detail. We have to be careful coming across as
%though we have written a paper ``not for a mainstearm audience'', but we can decline some of these requests
%at least for the main text) as not possible because of space constraints. This particular request does not 
%make perfect sense because of grammatic problems. But overall the referee seems to want us to explain the
%effect intuitively, whereas all we've really done is demonstrate its importance. If we want to insert an intuitive 
%explanation, that'd be more than we've tried for. We should talk about whether we have or could make
%something like that.
\noindent ***\emph{response} \\

@@

\noindent C) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system? this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones. \\

\noindent ***\emph{response} \\
%I think the earlier papers already showed that, at least for the parameters used there? Check. And the other parameters
%are experimentally manipulable, so this should be OK, but is worth discussing the extent to which ATAs were important
%for the original parameters. The it would be worth noting that the strongest ATA contributions come for parameters
%which were not imposed in the original experiment, so it may be worth repeating the experiment but with these other
%parameters.

@@

\noindent \textbf{Comments of the editor:} \\

\noindent C) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\

\noindent ***\emph{Thank you for pointing out the novelty.} \\

@@

\noindent C) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists. My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing. The authors have to make an effort to use a more biological/ecological language with reference to real world potential examples/cases/systems. Some mathematical material could be moved to Supplementary material. \\

%Need to address. However, this papers seems no more technical than the earlier Ellner paper. And, furthermore, 
%one of the referees (ref 1) actually asked us to put BACK technical content in the main text which we had 
%moved to the SI before submission to keep it less technical. That said, we do need to do a pass thu and 
%figure out what technical content we can reduce. 
\noindent ***\emph{response} \\

### 2023-07-03 paper/RefereeResponses1.tex

\noindent C2) Specifically, the focus is on asymmetric tail association (ATA), by which two random variables with given, fixed marginal distributions and correlation can nevertheless differ in the way that their right and left tails are correlated. This is one among many ways that one could characterize or decompose the full bivariate distribution between competition and environment, but the authors argue and demonstrate that ATAs of this kind are (a) prevalent in natural systems and (b) can have a meaningful and sometimes qualitative effect on the outcome of coexistence between two species.  The combination of results for theoretical distirbutions but also for experimental data is very powerful and convincing. \\

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C3) I have one overarching question, which I think may be important for ensuring that those points (a) and (b) are robust and then some minor comments that may improve the readability of the manuscript for the EL audience. \\

\noindent ***\emph{We address your questions and comments below.} \\

@@

\noindent C5) Perhaps equally important to clarifying this process for the readership is whether this process is unique. I could not quite get a sense of whether there could be alternative versions of this removal of ATA from a bivariate distribution, that still satisfy the authors' constraints on marginal distributions and overall correlation between the two variables. Are there alternatives, or is the authors' procedure unique in achieving the desired outcome?  If the former, then I would want to understand if the outcomes downstream tend to be robust to the specific method used. If the latter, it would be helpful to have an explanation as to why this is a unique decomposition. \\

\noindent ***\emph{<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>} \\
%Dan to do

\noindent C6) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
$E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.  Maybe this is standard notation, but somehow it may be clearer to put the subscript on the E for expectation value, to indicate that the probability distribution used to generate the definition of the averaging is what differs between the two expressions, while the variables E and C themselves are in the end just dummy variables to be integrated over.  \\

\noindent ***\emph{<This one asks for possibly changing notation. I don’t think we should do it, but we need to explain in a response what is actually going on, why we chose the notation that we do, and then add a new SM section addressing what the referee is talking about. Actually it’s the bivariate distribution that changes when you apply ||, tho the marginals are not changed. So the referee does have a bit of a point – E^{||} by itself is actually the same distribution as E, it’s just that it’s a different random variable in the sense of being related differently to C^{||} than E is to C. This is the same issue with the # notation that has already been established - elaborate. It’s actually formally correct in the sense of random variables if we view *all* our random variables as being measurable functions from the *same* probability space. So in the end the referee does not really have a point, the notation is formally correct. But you can explain all this in a new sup mat section, and that's worth doing in its own right as well as to satisfy them and people like them.>} \\
%Dan to do

@@

\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).  These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species. Is this really a demonstration that ATAs occur for C and E?  If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).  Maybe I am missing something. \\

\noindent ***\emph{<Yes, the referee is right about this, that's what we're doing. Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it. But if we change the example, as suggetsed by another reviewer Q, we should mention that here, too.>} \\

@@

\noindent C) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations. Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics. I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript. \\

\noindent ***\emph{We have addressed the referee's specific suggestions below. We have also redoubled our
efforts to make the manuscript as accessible as possible.} \\
%We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
%sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
%area in ecology?

### 2023-07-06 paper/RefereeResponses1.tex

\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).  These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species. Is this really a demonstration that ATAs occur for C and E?  If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).  Maybe I am missing something. \\

\noindent ***\emph{<Yes, the referee is right about this, that's what we're doing. Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it. But if we change the example, as suggetsed by another reviewer Q, we should mention that here, too.>} \\

\noindent C8) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.  I think $r_{j \bs i}$ must also assume that $j$ is at steady state, but this isn't mentioned. Perhaps that could be added for completeness. \\

\noindent ***\emph{This was done already, the referee appears to have missed it. It was line 136 in the original submission,
now line XXX in the resubmitted version.} \\
%Don't forget to add the new line numbers at the end.

@@

\noindent C9) E.g. line 100 needs removal of the word "are" or the addition of the word "and" somewhere to make sense. I'd consider going through the paper carefully for any typos or small corrections of this kind. \\

\noindent ***\emph{We changed the wording as requested, and went through the entire manuscript again looking for typos
and grammatical errors.} \\
%Jasmin to look at line 100 and revise as needed.

@@

\noindent C12) I like the conceptual idea of removing the long tail. \\

\noindent ***\emph{Thank you for the positive feedback.} \\

\noindent C13) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does. I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed. Additionally, it might be beneficial to indicate which distribution is then used to compute which part of the decomposition. For example, in Figure S1 d you show $E_i^{||}$, but from my understanding this distribution would be used to compute $E_i^{(EC)}$, correct? \\

\noindent ***\emph{<Clearly since two referees have now mentioned it we need to put some version of Fig S1 back into the
main text. The earlier referee wanted to see the whole figure brought back, but this referee wants to see just part, i.e.,
a simpler version, as specified. It's not clear to me how it would help the reader to just see the three panels the referee
suggests. Only 2 of the suggested panels are currently part of Fig. S1 - the suggestion of putting in the uncorrelated
distribution would be new. The current Fig. S1 serves the purpose of describing exactly the process of removing ATAs. 
The new suggested figure would not serve that purpose, it would instead maybe give the reader a sense for what these
distributions look like when manipulated in the ways we are talking about. But it's not real clear to me that much
intuition will be conveyed by doing that. One way to reply would be just to add a panel to the figure that shows 
the uncorrelated case, and move the resulting 5-panel figure back in. Then we'd explain that 1) the other referee asked
for the whole figure, 2) it's not clear to us that the simplified version serves any good purpose, it certainly 
does not serve the purpose of explaining the process of removing ATAs, but 3) we went ahead and added the uncorrelated
case anyway, since that's what was used previously, so it makes sense to have that point of comparison. If we do
this, Jasmin can add the new panel to the figure.>} \\
%Jasmin to make the ammended figure, Dan to move it and the associatde text bck to main text
%The other main thing Jasmin should do is finalize the repo - checkpoint issues.

@@

\noindent C17) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\

\noindent ***\emph{There is no space to intro these things, so we limited this text to just mentioning 
extinction risks and stability of ecosystem functioning. Whereas some ecologists may not have heard of 
Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
a bell for most ecologists.} \\

\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\

\noindent ***\emph{We spent 15-20 minutes or so trying to add an explanation that was succinct enough to not break
the flow of the paper while still being informative, but we could not do it. Any more than a short sentence would break the 
flow, and we felt that all the one-sentence explanations we could come up with were unclear. 
FYI, in a nutshell the idea is as follows. Assume that low values of an environmental variable reduce the 
growth rates of local populations and high values elevate those growth rates. In that case one may
expect that left-tail associations increase metapopulation extinction risks because then environmental
``catastrophes'' are spatially synchronized, presumably reducing the potential for rescue effects. Whereas
``bonanzas'' (extremely good conditions, with a lot of growth) are much less correlated. The case where
high values of the variable reduce the growth and low values elevate it is similar, but it's right-tail association
that elevates extinction risk. This is most of the explanation, but it gets a bit more nuanced, still - 
see the paper for additional details. Thanks for your interest, by the way.} \\
%DAN: maybe add to the above to point out that they don't need to know this to understand what we're doing in this paper!

@@

\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\

\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
and main ideas of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
defined in Theory, which they are.} \\

@@

\noindent C22) L74: The optimum might actually be the same, but they need different responses. \\

\noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\

\noindent C23) Eq. 1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\

\noindent ***\emph{We disagree, and would prefer not to make this change, for several reasons. 
First, the lottery model is a very simple model, at the same level of complexity
as models introduced in essentially all introductory 
undergraduate ecology courses (e.g., the Lotka-Volterra competition model), so there is no argument that 
the readership of Ecology Letters cannot easily understand the material presented here.
Second, this material serves a clear motivating purpose, most appropriately placed in the Introduction. 
The last author (DCR) has found, from decades of pitching 
quantitative papers to general ecological audiences, that clear motivation is essential - it is crucial to demonstrate 
clearly that an effect one plans to elaborate is present, and potentially important, prior to asking the reader to engage 
the kind of mental energy necessary to read the guts of the paper. So this 
paragraph is for showing that the effect of ATAs on coexistence can be real, at least for a simple 
example. That goal is convincingly 
accomplished by the paragraph. Given that the math is simple, and also more precise and definitive than 
qualitative verbal arguments that usually take up space in Introductions, we think this material is actually
more beneficial to the readership and more effective for the intended motivational purpose than would be more words
(and we see based on C25 below that the referee appears to agree with us here). 
Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
to want to exclude mathematics of any kind from the Introduction, especially when the math is simple,
accessible, and serves a clear motivational purpose, seems like a dated bias to us. 
Finally, we firmly believe that ecology as a discipline gets stronger as it continues the clear
trend of the last few decades of incorporating more quantitative approaches. We believe we
benefit the field by challenging assumptions and biases such as categorically not putting equations 
in the Introduction.
Overall, we think making this suggested change would make the paper weaker, and would pander to an
outmoded and unhelpful tradition. So we will not make the change unless the 
editor insists on it as a condition of publication.} \\
%Jasmin to read the above and let Dan know if it's convincing or not, and how it can be improved.

@@

\noindent C24) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\

\noindent ***\emph{OK, we made this change, even though the same equation already appears just a few lines
above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\
%Are we being too petulant, should we just do it and say we did it?

@@

\noindent C) L126: Do you do this in the appendix? \\

\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
To avoid making claims that we have not fully supported
with rigorous mathematics or other formal methods, we rephrased to say the iid assumption is 
``probably not necessary''. Thanks for spurring us to do this.} \\

\noindent C) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\

\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
the $\#$ term because that terms involves both of them varying (though independently).} \\
%Jasmin to check and make sure Dan is right , here. Maybe there is a better place to refer them to to support our claim?

@@

\noindent C) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\

\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
to do this the first time we use the hard brackets. So thanks for pointing this out.
We do not know what the referee means by ``using curly under brackets'' so have not done anything
in response to that part of this comment.} \\

\noindent C) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work (and potential for error) 
of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
paper. And we could instead spend that time working on the concepts underlying the next paper.} \\
%Jasmin to consider (don't do yet) how hard/error prone this change would be, taking into account the 
%formulas in the SI? Unless this is MUCH easier and MUCH less error prone than Dan's hunch, I suggest 
%we don't do it.

@@

\noindent C) L316: So why do you show values of $\sigma >6$? And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA. In 1d it simply appears that the variance scales with the density. 1e is quite unconvincing to me. That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example? \\

\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
trying to do here is accurately portray our results about
when ATAs were important and when they were not. To us, the most forthright way of doing that is to
show \emph{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
our results seems important when such is our goal. Furthermore, there's no real cost to showing
all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
matters for continuing to explore when ATAs are important and when not is probably the strength
of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
based on such a thing.

Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
%the original paper and extract from p-value info, or we could come up with a better example,
%or something else, not sure, let's discuss. Maybe also augment the figure to make it clearer?
%How about if Jasmin: 1) adds statistical support to caption; 2) adds to the figure some means of
%hopefully drawing people's attention to the right features of this thing.
} \\

@@

\noindent C) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations. Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics. I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript. \\

\noindent ***\emph{We have addressed the referee's specific suggestions below. We have also redoubled our
efforts to make the manuscript as accessible as possible.} \\
%We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
%sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
%area in ecology?

\noindent C) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences. \\

\noindent ***\emph{response} \\
%Jasmin, let's discuss. I cannot, myself, say, even now, precisely *how* or *why* ATAs promote or hinder coexistence. 
%In my view, the paper is about showing that they CAN promote or hinder coexistence, and to what extent, so that 
%future work can continue to explore when they do and when they don't, and precisely how. So I want to write a response
%to that effect. However, perhaps you have an intuition, which can be simply explained, as to how ATAs
%either promote or hinder coexistence. I know that's maybe a lot to ask, but if we have such a thing we should consider
%putting it in. Ecologists always want a two-sentence, math-less explanation for everything. If quantum mechanics were
%part of ecology they'd want a two-sentence intuitive explanation for it, even if it was meaningless aside making
%the reader feel better about themselves. Sometimes such a thing does not exist, but when it does exist it can,
%at least sometimes, be useful. For the record, I think most intuitive, verbal explanations of storage effects
%are mostly useless, but perhaps we can do better than those.
%
%Dan and Jasmin to discuss further. Jasmin to dig out old paragraph of text. Jasmin to work on this.

\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ...

\noindent C) ...but how ATAs enter in play is not well solved. It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented. \\

\noindent ***\emph{response} \\
%Same comment again, Jasmin can work on it.

@@

\noindent ***\emph{response} \\
%Same comment again, Jasmin can work on it.

\noindent C) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor. Could the authors briefly indicate how this could occur? \\

\noindent ***\emph{<It cannot. ATAs can only influence storage effects, not the other mechanisms, as explained on 
lines xyz. Is that enough? Elaborate and re-iterate stuff from the paper.>} \\
%Dan to do

@@

\noindent C) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\

\noindent ***\emph{<This is a good question. We need to find a place where the models are presented to motivate
them, AND a place later, in the Discussion (or a new SI section called ``Additional discussion points'') where
we talk about why ATAs matter for one but not the other. I think this is just because the
effects of extremes are accentuated in the beta one, and so the effects of associations in the extremes are
accentuated. But that's not unreasonable, since many species have environments which fluctuate between
being suitable and unsuitable.>} \\
%Jasmin to consider this while brainstorming the prior comment.

@@

\noindent C) The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model. \\

%This referee is asking us to explain a lot of things in more detail. We have to be careful coming across as
%though we have written a paper ``not for a mainstearm audience'', but we can decline some of these requests
%at least for the main text) as not possible because of space constraints. This particular request does not 
%make perfect sense because of grammatic problems. But overall the referee seems to want us to explain the
%effect intuitively, whereas all we've really done is demonstrate its importance. If we want to insert an intuitive 
%explanation, that'd be more than we've tried for. We should talk about whether we have or could make
%something like that.
\noindent ***\emph{response} \\
%Jasmin to also lump this in with her thinking about the earlier ref 3 comments. 
%Q: When you look at E and C marginals for the diatom model, do they have most of their mass in the tails, like the 
%beta-fecundidies lottery model, or most of their mass in the middle, like the normal-fecundities lottery model?

### 2023-07-10 paper/ExplanationOfATAEffectsWorkingDoc.tex
additional comments and generalization

@@

We mainly consider the lottery model. Though we begin with some more general notation, we quickly make an
assumption that I think will only hold for the lottery model.
We assume, as usual, that for whatever model we consider (a 2-species model), we can write $r_i$ as
a function of $E_i$ and $C_i$, and $r_i$ is an increasing function of $E_i$ and decreasing function
of $C_i$. Then ATA effects are defined as
\begin{align}
\Delta_i^{[EC]} &= (\olr_{i \bs i} - \olr_{j \bs i})-(\olrn_{i \bs i} - \olrn_{j \bs i}) \label{eq:ATAeffects1}\\
&= (\olr_{i \bs i} - \olrn_{i \bs i}) - (\olr_{j \bs i} - \olrn_{j \bs i}) \label{eq:ATAeffects2}\\
&= (\olr_{i \bs i} - \olrn_{i \bs i}).
\end{align}
The third term in (\ref{eq:ATAeffects2}) is $0$ because that represents growth of $j$ in a steady-state scenario,
and the fourth term is $0$, for the lottery model, because of reasoning in S5.1. So here is where we are 
getting specific to the lottery model, as warned above. DAN: Maybe the next iteration of this thinking should 
consider what happens when that fourth term does not vanish. Can we generalize?

@@

Viewing $\Delta_i^{[EC]}$ as an average of a function over a set of points distributed in a certain way
(with ATAs) minus an average of the same function over a set of points distributed in another way
(with ATAs removed) makes it clear why we would expect to commonly see substantial ATA effects 
on coexistence: removing ATAs really changes the way points in a bivairate distribution are 
distributed, in a way that can interact strongly with averaging over a function representing growth of
a potential invader. In fact, it seems clear that only very special functions will give the same results 
when averaged over a typical distribution with ATAs or with ATAs removed.

@@

>>>>

\section{Initial thinking}

@@

We mainly consider the lottery model. Though we begin with some more general notation, we quickly make an
assumption that I think will only hold for the lottery model.
We assume, as usual, that for whatever model we consider (a 2-species model), we can write $r_i$ as
a function of $E_i$ and $C_i$, and $r_i$ is an increasing function of $E_i$ and decreasing function
of $C_i$. Then ATA effects are defined as
\begin{align}
\Delta_i^{[EC]} &= (\olr_{i \bs i} - \olr_{j \bs i})-(\olrn_{i \bs i} - \olrn_{j \bs i}) \label{eq:ATAeffects1}\\
&= (\olr_{i \bs i} - \olrn_{i \bs i}) - (\olr_{j \bs i} - \olrn_{j \bs i}) \label{eq:ATAeffects2}\\
&= (\olr_{i \bs i} - \olrn_{i \bs i}).
\end{align}
The third term in (\ref{eq:ATAeffects2}) is $0$ because that represents growth of $j$ in a steady-state scenario,
and the fourth term is $0$, for the lottery model, because of reasoning in S5.1. So here is where we are 
getting specific to the lottery model, as warned above. Maybe the next iteration of this thinking should 
consider what happens when that fourth term does not vanish. Can we generalize? I think we can - see below for
some efforts to generalize which were added later.

@@

Viewing $\Delta_i^{[EC]}$ as an average of a function over a set of points distributed in a certain way
(with ATAs) minus an average of the same function over a set of points distributed in another way
(with ATAs removed) makes it clear why we would expect to commonly see substantial ATA effects 
on coexistence: removing ATAs really changes the way points in a bivairate distribution are 
distributed, in a way that can interact strongly with averaging over a function representing growth of
a potential invader. In fact, it seems clear that only very special functions will give the exact same results 
when averaged over a typical distribution with ATAs or with ATAs removed.

@@

\section{Efforts to generalize}

Equation (\ref{eq:ATAeffects2}) is what we had prior to specifying to the lottery model. 
Above, we've considered the first two terms, which are the average of $r_i(E_i,C_i)$ over 
the distribution $(E_i,C_{i \bs i})$ minus the average of the same function
over $(E_i^\nat,C_{i \bs i}^\nat)$. But the last two terms of (\ref{eq:ATAeffects2})
are the average of $r_i(E_j,C_j)$ over the distribution $(E_j,C_{j \bs i})$ minus the 
average of the same function over $(E_j^\nat,C_{j \bs i}^\nat)$. So the same basic
ideas apply here as in the lottery model case, it's just that we have to make two comparisons 
of growth functions averaged over distributions, one comparison for the invader, $i$,
and the other for the resident, $j$. Again, essentially any model will have 
non-zero value of (\ref{eq:ATAeffects2}) -- only models designed as counterexamples
should typically have a precisely zero value for that expression. So ATAs will always
matter to some extent for coexistence. It's just a question of how much, for real systems. 
And, as far as can tell, that can only be answered by computing these quantities for
several real systems. We've shown that, at least for some systems, the quantities 
can be meaningfully large.

\section{Additional commentary/thinking by writing}

OK, it seems clear to me that basically any meanigful change to the distribution 
$(E,C)$ will typically alter the average value of the nonlinear function $r(E,C)$ 
over that distribution. There are many ways to decompose the growth rate 
$\olr_{i \bs i}$ through a series of alterations of the distribution $(E,C)$. 
Ellner et al. also make this point in their 2019 paper. The original Chesson 
theory is probably the most natural decomposition under a weak noise assumption,
but Ellner et al. and our paper make it clear there are other ways that may make
more sense, in some circumstances, if noise is not weak.
